---
title: "Concentrating or Relaxing: How My Activities Influence My Jazz Habits"
output: 
  flexdashboard::flex_dashboard:
    
    storyboard: true
    css: style.css
    theme: journal
  html_document: 
    keep_md: true 
---



```{r}
library(spotifyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(flexdashboard)

```

```{r, echo=FALSE}
genjazz <- get_playlist_audio_features("", "5vXOWi2SAMHjQUv2p77wok")
elejazz <- get_playlist_audio_features("", "2F2bLwuWwWWNwWb0VtZ0Ig")
jazz <- rbind(genjazz, elejazz)
```

### My Jazz Habits


```{r}


```

***

When I am studying or just relaxing I often listen to Jazz. It can set a mood and is often good background music whilst focussing on other things. I have got two main spotify jazz lists I listen to.
One of the lists is an arbitrary mix of different Jazz songs I enjoy. The songs are from different sub-genres and were composed at different times. 
The other jazz list consists of more "toned-down" versions of popular jazz songs. There are no vocals and abrupt changes in the song feel less intense. I call this my "elevator music"-jazz list.

There is a slight overlap between the two lists. For example, the classic jazz song "Take five - The Dave Bruback Quartet" is part of general jazz list, whilst a different version "Take 5 - New York Jazz Lounge" is part of the elevator jazz list. These songs are extremely similar, yet sometimes I want to listen to the normal jazz list and sometimes to the elevator jazz list.

This leads into multiple interesting things that can be researched:

- How do these list differ from each other and why?

- Can this explain my decisions in when to listen to which playlist?

- What similarities can be found throughout the whole corpus?

### Exploring the Playlists

```{r, echo=FALSE}

p1 <- ggplot(jazz, aes(x=playlist_name, y = energy, fill = playlist_name)) +
  geom_boxplot() + 
  theme(legend.position = "none", axis.title.x = element_blank())


p2 <- ggplot(jazz, aes(x=playlist_name, y =acousticness, fill = playlist_name)) +
  geom_boxplot() + 
  theme(legend.position = "none", axis.title.x = element_blank())


p3 <- ggplot(jazz, aes(x=playlist_name, y = speechiness, fill = playlist_name)) +
  geom_boxplot() + 
  theme(legend.position = "none", axis.title.x = element_blank())+ 
  scale_y_log10()
  
p4 <-ggplot(jazz, aes(x=playlist_name, y = danceability, fill = playlist_name)) +
  geom_boxplot() + 
  theme(legend.position = "none", axis.title.x = element_blank())

grid.arrange(p1, p2, p3, p4, ncol=2)

```

***


### Further Evaluation

```{r, echo=FALSE}

ggplot(jazz, aes(x=energy, y=acousticness, color=playlist_name)) +
  geom_point() 

```


***


### Conclusion

```{r}

```
